---
output: pdf_document
---
# Reproducible Research: Peer Assessment 1

```{r, setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, result="hide", fig.width=11)
library(reshape2)
library(lattice)
library(ggplot2)
library(tidyr)
library(dplyr)
# library(plyr)
```
## Loading and preprocessing the data

```{r}
 download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "rawdata.zip")
unzip("rawdata.zip")
data<- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

### Make a histogram of the total number of steps taken each day.

```{r}
data<- read.csv("activity.csv")
dat <- select(data, steps, date) %>%
    group_by(date) %>% 
    summarize(total_steps= sum(steps, na.rm = T)) 
    with(dat, hist(total_steps, main="figures/Total number of steps taken each day"))
    dev.copy(png,"plot1.png")
    dev.off()
```

### Report the mean and median total number of steps taken per day.

The mean and median total number of steps taken per day:

```{r}
select(dat, total_steps) %>%
summarise(mean= mean(total_steps), median= median(total_steps)) %>% as.data.frame
```

## What is the average daily activity pattern?

### Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r} 
library(lubridate)  
dat <- group_by(data, interval) %>%
    summarize(mean_steps= mean(steps, na.rm = T)) %>%
    with(plot(interval, mean_steps,main = "Average Number of Steps Taken Each Day", type = 'l', xlab = "5-min Interval", ylab = "Average Number of Steps"))
    
    dev.copy(png,"plot2.png")
    dev.off()
```

### The 5-minute interval contains the maximum number of steps

```{r}
    group_by(data, interval) %>%
    summarize(mean_steps= mean(steps, na.rm = T)) %>%
    filter(mean_steps==max(mean_steps), na.rm = T) %>%
    select(the_interval_maximize_steps= interval) %>% as.data.frame
```

## Imputing missing values

### The total number of rows with missing data is:

```{r}
sum(is.na(data$steps))
```

### The strategy for filling in all of the missing values is to replace the missing values with the mean for that 5-minute interval. 

```{r}
## mean steps table for intervals
dat <- group_by(data, interval) %>%
    summarize(mean_steps= mean(steps, na.rm = T)) %>% as.data.frame

## the replace function to replace the NAs with the means steps across days
replace_value <- function(steps,interval_value){
    if (!is.na(steps)) {steps}
    else {dat[dat$interval==interval_value,2]}
}  

## replace the NAs in the data
n= length(data$steps)
for (i in 1:n){ 
    data$steps[i]= replace_value(data$steps[i],data$interval[i])
}
```

### analysis with the new data

```{r}
dat <- select(data, steps, date) %>%
    group_by(date) %>% 
    summarize(total_steps= sum(steps))
    with(dat, hist(total_steps, main="Total number of steps taken each day"))
    dev.copy(png,"plot3.png")
    dev.off()
```


The mean and median total number of steps taken per day.

```{r}
select(dat, total_steps) %>%
summarise(mean= mean(total_steps), median= median(total_steps)) %>% as.data.frame
```

From the report we can see that both values are increased. and the histigram becomes more concentrated to the central.

## Are there differences in activity patterns between weekdays and weekends?

### create a new vairable indicating the type of days

```{r}
data<- mutate(data, date= as.Date(date))
day_type<- function(date){
        if (weekdays(date)=="Sunday"|weekdays(date)=="Saturday") {"weekend"}
        else {"workday"}
}

dat <- mutate(data, type= sapply(date,day_type)) %>%
    mutate(type = as.factor(type))
```

### From the following pictures we can see that the partterns are different.

```{r}
plotdat <- select(dat, steps, interval, type) %>%
    group_by(interval,type) %>%   
    summarize(steps= mean(steps)) 
    
    xyplot(steps-mean(steps)~interval|type, layout= c(1,2), data = plotdat, type ="l")
    dev.copy(png,"plot4.png")
    dev.off()
    qplot(interval, steps, data = plotdat, facets = .~type, geom = c("point","smooth"),method="loess")
    dev.copy(png,"plot5.png")
    dev.off()
```